---
title: "Data exploration"
output: github_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
```

# Loading data

```{r ilposdata}

load("../data/ilposdata.RData")
# load("../data/ilposdata.Rdata")

load("../data/paavodata.RData")

rename(ilposdata, sex= gender ) 

rm(aluejakokartat,cc,Data,hoobee_data,nb_donations_data,paavo,paavo_shares,paavo_vars,paavo18,preprocessed_paavo_data,summarised_donor_data,zipcode_maps,api_key,columns_to_mutate,corvallis,file,hb_data,helsinki,i,results,share_column_suffix,sp.vaesto,collapse_names,data,get_geo,map_fi_zipcode,map_fi_zipcode_interactive,order_columns,paavo_aggr,sum_finite,wmean,zip_code_map,fi_commune_number2name)
 
 ilposdata %>% mutate(zip = recode(zip,
                            "106001" = "10600",
                            "201001" = "20100",
                            "216001"=  "21600",
                            "257001" = "25710", 
                             "27001" = "27100",
                             "41301"= "41310" ,
                             "61001"="61100",
                             "651001"= "61100",
                             "669001"="66900",
                             "685001" = "68500",
                             "686001"= "68600",
                             "686201"= "68620",
                             "651001"= "06500",
                             "651001" = "65100",
                             "652001" = "65200",
                              "669001"= "66900",
                              "686001" = "68600",
                              "688001" = "68600", 
                               "688001" = "68600")) 

```

# table 1 
```{r}


#table 1 
library(arsenal)

tab1 <- tableby(gender ~ Hb + age+ aborh + age.group + FirstEvent, data=ilposdata)
summary(tab1, text=T)
results="asis"

# remove p-values + missing values and make it look prettier
```


# Preprocessing

You can also embed plots, for example:
## Donation data
```{r pressure, echo=TRUE}
nb_donations_data <-
  ilposdata %>% 
    mutate(Year = year(dateonly)) %>% 
    filter(donat_phleb == "K") %>% 
    count(donor,Year, zip)
 
 summarised_donor_data <- 
  ilposdata %>% 
    mutate(Year = year(dateonly),
           Hb = ifelse(is.nan(Hb), NA, Hb)) %>% 
    group_by(donor, Year, zip) %>% 
    summarize(mean_hb=mean(Hb, na.rm = TRUE),
              aborh=last(aborh),
              gender=last(gender),
              age=last(age)
                           ) %>% 
   drop_na(mean_hb) %>% 
   full_join(nb_donations_data,
             by = c("donor", "zip","Year")) %>% 
   mutate(n = replace_na(n, 0))

```

## Paavo data
```{r}
preprocessed_paavo_data <- 
  paavodata %>% 
    dplyr::select(pono, vuosi, hr_mtu, hr_ktu,nimi, ko_perus, ko_yliop, ko_ammat, ko_al_kork, ko_yl_kork, pt_tyott, pt_opisk, pt_tyoll, te_omis_as, te_vuok_as,he_vakiy, ko_ika18y) %>% 
    rename(zip = pono,
           Year= vuosi,
           Basiceducation=ko_perus,
           unemployed = pt_tyott,
           students = pt_opisk,
           employed = pt_tyoll,
           owner_apartment = te_omis_as,
           rental_apartment = te_vuok_as,
           medianincome= hr_mtu,
           averageincome= hr_ktu,
           population= he_vakiy,
           population18= ko_ika18y)

#Remove the old variables, which were combined
preprocessed_paavo_data["secondaryeducation"] <- rowSums(preprocessed_paavo_data[c("ko_yliop",
                                  "ko_ammat")])
preprocessed_paavo_data["tertiaryeducation"] <- rowSums(preprocessed_paavo_data[c("ko_al_kork",
                                 "ko_yl_kork")])

preprocessed_paavo_data <- subset(preprocessed_paavo_data, select=-c(ko_yliop,
                                                           ko_ammat, ko_al_kork,
                                                                     ko_yl_kork))




final_data <-
  left_join(summarised_donor_data,
            preprocessed_paavo_data,
            by = c("zip", "Year")) %>% 
  filter(Year > 2014)
```

# Preliminary plots

```{r}
final_data %>% 
  ggplot(aes(x = medianincome)) +
  geom_histogram(binwidth = 1000) +
  facet_grid(Year ~. )


final_data %>% 
  ggplot(aes(x = averageincome)) +
  geom_histogram(binwidth = 1000) +
  facet_grid(Year ~. )
```


```{r}
final_data %>% 
  ggplot(aes(x = averageincome)) +
  geom_histogram(binwidth = 1000) +
  facet_grid(Year ~. ) +
  scale_y_log10()
```

```{r}
final_data %>% 
  ggplot(aes(x = Year, y = averageincome)) +
  geom_jitter(alpha = 0.2)
```





```{r} 
final_data %>% 
ggplot(aes(x = Basiceducation)) +
geom_histogram(binwidth = 1000) +
  facet_grid(Year ~. )


```
```{r}
final_data %>% 
ggplot(aes(x = secondaryeducation)) +
geom_histogram(binwidth = 1000) +
  facet_grid(Year ~. )

```

```{r}
final_data %>% 
ggplot(aes(x = tertiaryeducation)) +
geom_histogram(binwidth = 1000) +
  facet_grid(Year ~. )
```

# Plotting housing variables 
```{r}
final_data  %>% 
ggplot(aes(x = owner_apartment)) +
geom_histogram(binwidth = 1000) +
  facet_grid(Year ~. )

final_data  %>% 
ggplot(aes(x = rental_apartment)) +
geom_histogram(binwidth = 1000) +
  facet_grid(Year ~. )

```

# Plotting main type of activity 
```{r}
final_data  %>% 
ggplot(aes(x = unemployed)) +
geom_histogram(binwidth = 1000) +
  facet_grid(Year ~. )

final_data  %>% 
ggplot(aes(x = employed)) +
geom_histogram(binwidth = 1000) +
  facet_grid(Year ~. )

final_data  %>% 
ggplot(aes(x = students)) +
geom_histogram(binwidth = 1000) +
  facet_grid(Year ~. )


```


  
# summarised by hb 
```{r}

hb_by_zip <- final_data %>% 
  group_by(zip, Year) %>% 
  summarise(mean_hb=mean(mean_hb)) 

hb_by_zip %>% 
  ggplot(aes(x = Year, y = mean_hb)) +
  geom_jitter(alpha = 0.2)


```






# correlation matrix and plots
#Paavodata 
```{r}
library("GGally")


varspaavo <- c("medianincome","averageincome","tertiaryeducation","secondaryeducation","rental_apartment", "employed","unemployed","Basiceducation")

ggpairs(final_data,columns= varspaavo)



```
#correlation matrix with donor data 

```{r}
library(GGally)
donorvars <- c ("Hb","age","gender","aborh")
ggpairs(ilposdata, columns= donorvars)
```
#ggpairs with both datasets
```{r}
varboth <-  c( "mean_hb", "tertiaryeducation","medianincome","unemployed")
ggpairs(final_data, columns= varboth)

```


Correlation matrix and  plots on final data with psycho package
```{r}
library(psycho)

 
cor <- correlation (final_data) 


plot(cor)
summary(cor)
```




# Count division of post codes between mobile and fixed sites.  
```{r}

    Sitedata <- ilposdata %>%
    mutate(locSim = case_when(
      Site == "H0091" | Site == "H0092" ~ "Kivihaka",
      Site == "H0093"  ~ "Sanomatalo",
      Site == "H0096"  ~ "Espoo",
      Site == "K0297" ~ "Kuopio",
      Site == "T0837" ~ "Tampere",
      Site == "T07343" ~ "Seinäjoki",
      Site == "T0179" ~ "Jyväskylä",
      Site == "T0398" ~ "Lahti",
      Site == "L0564" ~ "Oulu",
       Site == "U0853" ~ "Turku",
      TRUE ~ "mobilesite")) %>% 
      summarise(fixedsite 
      

  
 Sitedata %>% 
mutate(Year = year(dateonly)) %>% 
filter(donat_phleb == "K") %>% 
filter(Year > 2014) %>% 
count(locSim, Year) 

   
 # make a new variable (fixedsites or plot without mobilesites)
 


```


 
```{r}

prop_donor <- ilposdata %>% 
  mutate(Year = year(dateonly)) %>% 
  filter(donat_phleb == "K") %>% 
    count(donor, Year,zip) %>% 
    count(Year, zip) %>% 
    filter(Year == 2017 | Year == 2018) %>%
    rename(nb_donors = n)

prop_paavo <- paavodata %>% 
  rename(zip = pono, Year= vuosi) %>%   
  filter(Year == 2019) %>% 
  mutate(eligible_population = he_18_19+ he_20_24+ he_25_29+ he_30_34+ he_40_44+
                                  he_45_49+ he_50_54+ he_55_59+ he_60_64+ he_65_69) %>% 
  dplyr::select(-Year) %>% 
  full_join(prop_donor,by = c("zip")) %>% 
  dplyr::select(zip, Year, eligible_population, nb_donors, hr_mtu, hr_ktu,nimi, ko_perus, 
                ko_yliop, ko_ammat, ko_al_kork, ko_yl_kork, pt_tyott, pt_opisk, pt_tyoll,
                te_omis_as, te_vuok_as,he_vakiy,  ko_ika18y, hr_tuy) %>% 
    rename(Basiceducation=ko_perus,
           unemployed = pt_tyott,
           students = pt_opisk,
           employed = pt_tyoll,
           owner_apartment = te_omis_as,
           rental_apartment = te_vuok_as,
           medianincome= hr_mtu,
           averageincome= hr_ktu,
           population= he_vakiy,
           population18= ko_ika18y)%>%
mutate(prop_donor= nb_donors/eligible_population, 
      prop_basic=Basiceducation/hr_tuy)



#lets look at  missing values after recoding few zips 
#prop_paavo %>% filter(is.na(eligible_population)) %>% summarise(total = sum(nb_donors, na.rm = TRUE))
#  total=   8884

```
 
```{r}
prop_paavo %>% 
  filter(prop_basic <= 1) %>% 
  ggplot(aes(x = prop_donor, y = prop_basic)) +
  geom_point(alpha = 0.25) +
  facet_grid(Year ~.)



```
 
 
```{r}
prop_paavo %>% 
  filter(prop_basic > 1)
```
 


```{r}

```

