---
title: "Visualisation"
output: github_document
---

```{r}
load("/home/ilpo/Paavo/data/preprocessed.2019-07-24.RData")

```


```{r}
knitr::opts_chunk$set(echo = TRUE)

```
```{r}
setwd(dir="/home/ilpo/Paavo/src")
getwd()

```



```{r cars}

library(tidyverse)
library(lubridate)
library(viridis)




```

#Relationship between proportion of donors and median income. 

```{r}
df <- as.tbl(preprosessed_paavo) %>% filter(!is.na(prop_donors))
p <- ggplot(data=df)
p <- p +  geom_point(mapping=aes(y = prop_donors, x = medianincome, colour=eligible_population))
p <- p +   scale_x_log10()+
  geom_smooth(aes(y = prop_donors, x = medianincome)) +
 scale_color_viridis(discrete=FALSE,direction = -1,trans="log")+
labs(x = "Median income",
        y = "Proportion of donors",
        title = "Proportion of donors per median income of postal codes per year")+
facet_grid(Year ~.)
p
```
Size matters. Population and outliers needs to be controlled before any modeling.  ¤

##Filtered version 


```{r}
#filtered_paavo <- preprosessed_paavo %>% 
##ilter(eligible_population > 99) %>%
#filter(!is.na(prop_donors)) 


ggplot(data= filtered_paavo,mapping= aes(x= medianincome, y= prop_donors))+
geom_point(mapping=aes(y = prop_donors, x = medianincome, colour=eligible_population))+ 
  scale_x_log10()+
  geom_smooth(aes(y = prop_donors, x = medianincome)) +
 scale_color_viridis(discrete=FALSE,direction = -1,trans="log")+
labs(x = "Median income",
        y = "Proportion of donors",
        title = "Proportion of donors per median income of postal codes per year")+
facet_grid(Year ~.)
```


## New donors and per median income
```{r}
df <- as.tbl(preprosessed_paavo) %>% filter(!is.na(prop_new_donors))
p <- ggplot(data=df)
p <- p +  geom_point(mapping=aes(y = prop_new_donors, x = medianincome, colour=eligible_population))
p <- p +   scale_x_log10()+
  geom_smooth(aes(y = prop_new_donors, x = medianincome)) +
 scale_color_viridis(discrete=FALSE,direction = -1,trans="log")+
labs(x = "Median income",
        y = "Proportion of new donors",
        title = "Proportion of new donors per median income of postal codes per year")+
facet_grid(Year ~.)
p
```
New donors looks really weird. It could be by low numbers of first time donors in many postal code areas.  Maybe need to be modelled on their own or at least filtered. eg vähintään 100

##repeated donors 

```{r}
df <- as.tbl(preprosessed_paavo) %>% filter(!is.na(prop_new_donors))
p <- ggplot(data=df)
p <- p +  geom_point(mapping=aes(y = prop_new_donors, x = medianincome, colour=eligible_population))+
  scale_color_viridis(discrete=FALSE,direction = -1,trans="log")+
  facet_grid(Year~.)
p
```


```{r}
df <- as.tbl(preprosessed_paavo) %>% filter(!is.na(prop_donors))
p <- ggplot(data=df)
p <- p +  geom_point(mapping=aes(y = prop_repeat_donors, x = medianincome, colour=eligible_population))+
  scale_color_viridis(discrete=FALSE,direction = -1,trans="log")+
geom_smooth(aes(y = prop_repeat_donors, x = medianincome)) + 
facet_grid (Year~.)
p
```


```{r}
df <- as.tbl(preprosessed_paavo) %>% filter(!is.na(prop_repeat_donors))
p <- ggplot(data=df)
p <- p +  geom_point(mapping=aes(y = prop_repeat_donors, x = medianincome, colour=eligible_population))
p <- p +   scale_x_log10()+
  geom_smooth(aes(y = prop_repeat_donors, x = medianincome)) +
 scale_color_viridis(discrete=FALSE,direction = -1,trans="log")+
labs(x = "Median income",
        y = "Proportion of repeat donors",
        title = "Proportion of repeat donors per median income of postal codes per year")+
facet_grid(Year ~.)
p
```
Repeat donors also have high amount of outliers, where there are no postal codes with donors.  Kato onko repeat donoreita joka paikassa. Vertaa repeattia ja newta missä ei ole jompaa kumpaa. Suodatuksen jälkeen millaisista paikoista ei ole repeatteja. 

# Relationship between higher education and proportion of donors
```{r}

p <- ggplot(data=df)
p <- p +  geom_point(mapping=aes(y = prop_donors, x = proportion_inhabitants_with_higher_education, colour=eligible_population))
p <- p +   scale_x_log10() + geom_smooth(aes(y = prop_donors, x = proportion_inhabitants_with_higher_education)) +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
  facet_grid(Year~.)
p

```

Alle sataset pois eligiblesta 
filtered version = eligible population > 100

```{r}
ggplot(data= filtered_paavo, mapping=aes(y = prop_new_donors, x = proportion_inhabitants_with_higher_education, colour=eligible_population))+
geom_point(mapping=aes(y = prop_new_donors, x = proportion_inhabitants_with_higher_education, colour=eligible_population))+
  scale_x_log10() + 
geom_smooth(aes(y = prop_new_donors, x = proportion_inhabitants_with_higher_education)) +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") 
  #facet_grid(Year~.)
```

 
##First time donors
```{r}
p <- ggplot(data=df)
p <- p +  geom_point(mapping=aes(y = prop_new_donors, x = proportion_inhabitants_with_higher_education, colour=eligible_population))
p <- p +   scale_x_log10() + geom_smooth(aes(y = prop_new_donors, x = proportion_inhabitants_with_higher_education)) +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") +
  facet_grid(Year~.)
p
```



## repeated donors
```{r}
p <- ggplot(data=df)
p <- p +  geom_point(mapping=aes(y = prop_repeat_donors, x = proportion_inhabitants_with_higher_education, colour=eligible_population))
p <- p +   scale_x_log10() + 
geom_smooth(aes(y = prop_repeat_donors, x = proportion_inhabitants_with_higher_education)) +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") +
  facet_grid(Year~.)
p
```






```{r}
mörkö <- preprosessed_paavo %>% 
filter(prop_donors > 0.04) %>% 
filter(Year == 2018) %>% 
filter(eligible_population > 400)
ggplot(data=mörkö, mapping= aes(x= medianincome, y= prop_donors )) +
geom_point() +
geom_text(data = mörkö,check_overlap =TRUE,show.legend = TRUE,hjust = 0,
mapping = aes(label = nimi)) 


```
## First attempt to visualise areas with population size and proportion of donors.  It has raised a question that Finnish-Swedish population could have higher proportion of blood doning. Needs further investigating. Kuntatasolla tai postinumerotasolla suomenruotsalaisten alueet. 



```{r}


ggplot(data=preprosessed_paavo, mapping= aes(x= medianincome, y= prop_donors,  color=prop_repeat_donors))+
geom_point ()+ 
scale_x_log10() +
geom_smooth()+
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") +
facet_wrap (Year~.)


```
```{r}
ggplot(data=preprosessed_paavo, mapping= aes(x= medianincome, y= prop_donors,  color=prop_new_donors))+
geom_point ()+ 
scale_x_log10() +
geom_smooth()+
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") +
facet_wrap (Year~.)
```
# Mediaani ja keskiansio juttu jos löydät. 

#
```{r}
preprosessed_paavo %>% 
ggplot(aes(x= prop_donors)) +
geom_histogram() +
facet_grid(Year~.)

preprosessed_paavo %>% 
ggplot(aes(x= prop_donors)) +
geom_histogram() +
scale_x_log10() + 
facet_grid(Year~.)
```
```{r}
preprosessed_paavo %>% 
ggplot(aes(x= proportion_inhabitants_with_higher_education)) +
geom_histogram() +
scale_x_log10() + 
facet_grid(Year~.)

ggplot(data=preprosessed_paavo, mapping = aes(x= prop_repeat_donors))+
geom_histogram ()


ggplot(data=preprosessed_paavo, mapping = aes(x= prop_new_donors))+
geom_histogram ()


ggplot(data=preprosessed_paavo, mapping = aes(x= medianincome))+
geom_histogram () + 
scale_x_log10() + 
facet_grid(Year~.)


ggplot(data=preprosessed_paavo, mapping = aes(x= averageincome))+
geom_histogram () + 
scale_x_log10() + 
facet_grid(Year~.)

ggplot(data=preprosessed_paavo, mapping = aes(x= eligible_population))+
geom_histogram () 


```

