---
title: "Distance calculating"
author: "Ilpo Arminen"
date: "8/6/2019"
output: github_document
---


# This data is downloaded from geonames website.(*add blaablaa*)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
 df <- readr::read_tsv("coordinates.txt",col_names= FALSE)
load("../data/paavodata.rdata")
load("../data/ilposdata.rdata")
load("/home/ilpo/Paavo/data/site_codes.rdata")
```
#loadin the file 



```{r}
library(tidyverse)
library(geosphere)
```

```{r}
 df <- readr::read_tsv("coordinates.txt",col_names= FALSE)
# This is downloaded from geonames.com. It has all the zip codes of Finland with longitude and latitude. 

  Uusimaa_coordinates <- df %>% 
dplyr::select(-X1,-X5, -X7,-X9,-X12,-X6,-X8) %>%  # unnecessary variables
filter(X4 == "Uusimaa") %>%               #Filtered to uusimaa region
rename(zip=X2, city= X3, region=X4, longitude= X10, latitude= X11) #renaming 

```











## Example 
#The haversine formula determines the great-circle distance between two points on a sphere given their longitudes and latitudes. Important in navigation, it is a special case of a more general formula in spherical trigonometry, the law of haversines, that relates the sides and angles of spherical triangles.# Haversine Formula (from R.W. Sinnott, "Virtues of the Haversine", Sky and Telescope, vol. 68, no. 2, 1984, p. 159):

    dlon = lon2 - lon1
    dlat = lat2 - lat1
    a = sin^2(dlat/2) + cos(lat1) * cos(lat2) * sin^2(dlon/2)
    c = 2 * arcsin(min(1,sqrt(a)))
    d = R * c


```{r}
                  #  Kamppi              Punavuori
distHaversine(p1=c(60.1714, 24.9316), p2= c( 60.1632, 24.9391))

distVincentyEllipsoid(p1=c(60.1714, 24.9316), p2= c(60.1632, 24.9391))



                  #Kamppi                 Kiminkylä
distHaversine(p1= c(60.1714 , 24.9316), p2= c(60.7121 , 26.2904))
```
## Atleast distHaversine function seems to work allright. Good news... 


## Filtering the site_codes_data

```{r}
site_zip <- site_codes_data %>% 
select(c(1,2,5)) %>% #selecting only Site, site name and site zip. 
rename(Site=code, name=site_name, site_zip=zip)   # renamed to match variable names in processed paavo data 



```

```{r}


new_data <- left_join(processed_site_paavo, Uusimaa_coordinates, by= c("zip")) #seems to be ok. Now preprocessed file has coordinates

```



```{r}

#distm(c(lon1, lat1), c(lon2, lat2), fun = distHaversine)
#distHaversine()
#distMeeus()
#distRhumb()
#distVincentyEllipsoid()
#distVincentySphere()

#map(x=new_data,distVincentyEllipsoid(p1= 60.2106, 24.9035, p2= longitude, latitude))
  

```




```{r}
#Kivihaka long	60.2106	 lat 24.9035
```


```{r}
get_geo_distance = function(long1, lat1, long2, lat2, units = "miles") {
  loadNamespace("purrr")
  loadNamespace("geosphere")
  longlat1 = purrr::map2(long1, lat1, function(x,y) c(x,y))
  longlat2 = purrr::map2(long2, lat2, function(x,y) c(x,y))
  distance_list = purrr::map2(longlat1, longlat2, function(x,y) geosphere::distHaversine(x, y))
  distance_m = list_extract(distance_list, position = 1)
  if (units == "km") {
    distance = distance_m / 1000.0;
  }
  else if (units == "miles") {
    distance = distance_m / 1609.344
  }
  else {
    distance = distance_m
    # This will return in meter as same way as distHaversine function. 
  }
    distance
}  
```

# example but not really good one 

```{r}
list1 <- data.frame(longitude = c(80.15998, 72.89125, 77.65032, 77.60599, 
                                  72.88120, 76.65460, 72.88232, 77.49186, 
                                  72.82228, 72.88871), 
                    latitude = c(12.90524, 19.08120, 12.97238, 12.90927, 
                                 19.08225, 12.81447, 19.08241, 13.00984,
                                 18.99347, 19.07990))
list2 <- data.frame(longitude = c(72.89537, 77.65094, 73.95325, 72.96746, 
                                  77.65058, 77.66715, 77.64214, 77.58415,
                                  77.76180, 76.65460), 
                    latitude = c(19.07726, 13.03902, 18.50330, 19.16764, 
                                 12.90871, 13.01693, 13.00954, 12.92079,
                                 13.02212, 12.81447)) 
                   # locality = c("A", "A", "B", "B", "C", "C", "C", "D", "D", "E"))
                    
                    


# create distance matrix
mat <- distm(list1[,c('longitude','latitude')], list2[,c('longitude','latitude')], fun=distVincentyEllipsoid)

# assign the name to the point in list1 based on shortest distance in the matrix
#list1$locality <- list2$locality[max.col(-mat)]                    
                    
  mat                  
                    
```
# 
```{r}                          
                                #etu-töölö
list3 <- data.frame(longitude = c(60.2106),
                      latitude = c(24.9035))
                                #punavuori    #kaartinkaupunki #kaivopuisto
list4 <- data.frame(longitude = c(60.1632, 60.1645, 60.1578),
                  latitude= c(24.9391, 24.9487,24.9525))



mat2 <- distm(list3[,c('longitude','latitude')], list4[,c('longitude','latitude')], fun=distVincentyEllipsoid)
mat2
```




```{r}
earth.dist <- function (long1, lat1, long2, lat2)
{
rad <- pi/180
a1 <- lat1 * rad
a2 <- long1 * rad
b1 <- lat2 * rad
b2 <- long2 * rad
dlon <- b2 - a2
dlat <- b1 - a1
a <- (sin(dlat/2))^2 + cos(a1) * cos(b1) * (sin(dlon/2))^2
c <- 2 * atan2(sqrt(a), sqrt(1 - a))
R <- 6378.145
d <- R * c
return(d)
}
```

```{r}
teppo <-  pmap(new_data, ~earth.dist(long1=60.2106, lat1=24.9035, long2=new_data$longitude, lat2 =new_data$latitude )) 


as.data.frame.list(teppo)
teppo



```
#Calculating distance between two points using the distm function inside mutate
```{r}
jaffattulloojahelematpaukkuu <- new_data %>% 
rowwise () %>% 
  mutate(Dist = pmap_dfr(., ~
           distm(x = longitude, y = latitude, fun = distHaversine)))
```



