---

title: "distance_function"
author: "Ilpo Arminen"
date: "8/6/2019"
output: html_document
---


```{r}
 df <- readr::read_tsv("coordinates.txt",col_names= FALSE) # Getting the coordinates from geonames.org, because Geosphere package do not work with eurefin coordinates. 
#load("../data/ilposdata.rdata")
load("/home/ilpo/Paavo/data/site_codes.rdata")
load("/home/ilpo/Paavo/data/processed_site_paavo.RData") # This data is filtered by site 
```
#loadin the file 



```{r}
library(tidyverse)
library(geosphere)
library(table1)
```

```{r}
 df <- readr::read_tsv("coordinates.txt",col_names= FALSE)
# This is downloaded from geonames.com. It has all the zip codes of Finland with longitude and latitude. 

  Uusimaa_coordinates <- df %>% 
dplyr::select(-X1,-X5, -X7,-X9,-X12,-X6,-X8) %>%  # unnecessary variables
## filter(X4 == "Uusimaa") %>%           #Filtered to uusimaa region
rename(zip=X2, city= X3, region=X4, longitude= X10, latitude= X11) #renaming the variables 

```


## Filtering the site_codes_data

```{,recho=FALSE}
site_zip <- site_codes_data %>% 
select(c(1,2,5)) %>% #selecting only Site, site name and site zip. 
rename(Site=code, name=site_name, site_zip=zip)   # renamed to match variable names in processed paavo data 



```

```{r}


new_data <- left_join(processed_site_paavo, Uusimaa_coordinates, by= c("zip")) # Now preprocessed file has longitude and latitude instead of euref. 

```


                  #Kamppi                 Kiminkylä
distHaversine(p1= c(60.1714 , 24.9316), p2= c(60.7121 , 26.2904)) # There are quite a few different calculation to distance which all differs mostly on how round they assume that Earth is.  Couple of meters between Haversine and Vincenty ellipsoid doesnt really matter in this project. 
``` case of a more general formula in spherical trigonometry, the law of haversines, that relates the sides and angles of spherical triangles.# Haversine Formula (from R.W. Sinnott, "Virtues of the Haversine", Sky and Telescope, vol. 68, no. 2, 1984, p. 159):

    dlon = lon2 - lon1
    dlat = lat2 - lat1
    a = sin^2(dlat/2) + cos(lat1) * cos(lat2) * sin^2(dlon/2)
    c = 2 * arcsin(min(1,sqrt(a)))
    d = R * c
    
    
```{r}

###The haversine formula determines the great-circle distance between two points on a sphere given their longitudes and latitudes. VincentyEllipsoid differs on on Haversine on taking more exact picture on how round or ellipsoid the earth sphere is. 
                  #  Kamppi              Punavuori
distHaversine(p1=c(60.1714, 24.9316), p2= c( 60.1632, 24.9391))

distVincentyEllipsoid(p1=c(60.1714, 24.9316), p2= c(60.1632, 24.9391))

# Result in distance between Kamppi and punavuori differs about 2 meters, so its not gonna make a lot of bias on which to choose. 
```
## Calculating the distance between each fixed site and every zip, so that dataset is going to have 4 new variables which tells the distance to each fixed site.     

```{r}
final_data <- new_data %>% mutate(distKivihaka = distHaversine(cbind(60.2106, 24.9035), cbind(longitude,latitude))) %>% 
mutate(distEspoo = distHaversine(cbind(60.1525, 24.746), cbind(longitude,latitude))) %>% 
  mutate(distSanomatalo = distHaversine(cbind(60.1714, 24.9316), cbind(longitude,latitude))) %>% 
mutate(distLahti = distHaversine(cbind(60.9742, 25.6586), cbind(longitude,latitude)))
```



```{r, results="asis"}

label(final_data$medianincome)  <- "Median income of household per postal code"
label(final_data$averageincome) <- "Average income of household per postal code"
label(final_data$proportion_inhabitants_with_higher_education)  <- "Share of inhabitants with higher education per postal code"
label(final_data$prop_donors) <- "Proportion of donors per postal code "
label(final_data$nb_fixed_donations_per_zip) <- "Proportion of donations per postal code"
label(final_data$nb_fixed_donors_per_zip) <- "Number of donors per zip"
label(final_data$proportion_inhabitants_with_higher_education) <- "Proportion_inhabitants_with_higher_education"
label(final_data$nb_fixed_repeat_donors) <- "Number of donors who donated more than once per zip"
label(final_data$nb_fixed_first_time_donors) <- "Number of first time donors per zip"
label(final_data$nb_fixed_donation_per_act_donor) <- "Number of donations divided by numbers of donor per zip"
label(final_data$eligible_population) <- "Number of residents per zip who is suitable for donating blood"
units(final_data$medianincome) <- "Euro"
units(final_data$averageincome) <- "Euro" 

 table1(~prop_donors + eligible_population + nb_fixed_donors_per_zip + nb_fixed_donations_per_zip + nb_fixed_donors_per_zip + medianincome + averageincome + proportion_inhabitants_with_higher_education + proportion_inhabitants_with_higher_education + prop_repeat_donors + prop_new_donors + prop_donors + nb_fixed_donation_per_act_donor,data = final_data, overall = "Total", topclass="Rtable1-grid Rtable1-center")



``` 
137 206
## Calculating the min distance to the closest donation site. 

```{r}
 final_data$minDist <- apply(X = final_data[,c("distSanomatalo", "distKivihaka", "distEspoo", "distLahti")],FUN = min,MARGIN = 1,na.rm=TRUE)

```

### filtering the proportion of donors bigger than 99 for new dataset and removing Otaniemi as a outlier. Otaniemi has total of 173 donors and 286 donations. 
```{r}

mod_data <- final_data[!(final_data$zip %in% c("02150")), ]

model_data <- mod_data  %>% 
filter(nb_fixed_donations_per_zip > 99)

```

```{r}
save(model_data,
file = paste0("name.",Sys.Date(),".RData"))
```

