---
title: "Preprocessing fixed sites"
author: "Ilpo Arminen"
date: "7/31/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(dir ="/home/ilpo/Paavo/src")
load("../data/paavodata.rdata")
load("../data/ilposdata.rdata")
load("/home/ilpo/Paavo/data/site_codes.rdata")


```

```{r}
library(tidyverse)
library(lubridate)
```






# This is done by comparing missing values in zip codes with the list of Statistical centers Paavo data.
 #I took all the zip codes from blood doning data which had more than 50 missing values and  did not have match 
# in paavodata.

```{r}
ilposdata %>% mutate(zip = recode(zip,
#old value = new value
"106001" = "10600",
"201001" = "20100",
"216001"= "21600",
"257001" = "25710",
"27001" = "27100",
"41301"= "41310" ,
"61001"="61100",
"651001"= "61100",
"669001"="66900",
"685001" = "68500",
"686001"= "68600",
"686201"= "68620",
"651001"= "06500",
"651001" = "65100",
"652001" = "65200",
"669001"= "66900",
"686001" = "68600",
"688001" = "68600",
"688001" = "68600"))

```

# Removing string value "Na" which shows in plots. 
```{r}
ilposdata$zip <- gsub('.*NA.*',NA,ilposdata$zip)
ilposdata <- ilposdata %>% 
filter(!is.na(zip))
```



##Numbers of donors that donated to a fixed site per zip
```{r}
prepocessing <- ilposdata %>%
mutate(Year = year(dateonly)) %>% # Getting the variable Year from dates
filter(donat_phleb == "K") %>%    # filtering full blood donations only
filter(Year == 2017 | Year == 2018) %>%  # Selecting wanted years
 filter(Site %in% c("H0091", "H0092", "H0093","H0096","T0398")) %>%  # Selecting only Kivihaka,Espoo,Sanomatalo and Lahti fixed sites
count(donor, Year,zip) %>%   #  Selecting wanted variables
count(Year, zip) %>%        # Needed for counting donors per  zip codes and yearly count
rename(nb_fixed_donors_per_zip = n)  # renaming the n to numbers of donors per postal code

```


## number of donations that donated to a fixed site per
```{r}
test <-ilposdata %>%  #new subset
mutate(Year = year(dateonly)) %>%  # Getting the variable Year from dates
filter(donat_phleb == "K") %>%       # filtering full blood donations only
filter(Year == 2017| Year == 2018) %>%
filter(Site %in% c("H0091", "H0092", "H0093","H0096","T0398")) %>% # Selecting wanted years 
count(Year,zip) %>%   # Counting the donors straight to year and postal code. Not counting the donors means getting donations per postal code
rename(nb_fixed_donations_per_zip=n) 


```
## joining the data with commong variables

```{r}

preprosessing <-left_join(prepocessing,test,
by = c("zip", "Year"))
```

# Nb of first time donors that donated to a fixed site per
```{r}
firstevent <- ilposdata %>%
mutate(Year = year(dateonly)) %>%
filter(donat_phleb == "K") %>%
filter(Year == 2017| Year== 2018) %>%
 filter(Site %in% c("H0091", "H0092", "H0093","H0096","T0398")) %>% 
dplyr::select(zip,FirstEvent,Year) %>%
filter(FirstEvent == TRUE) %>%  #For selecting only donors who had their first blood donation in the selected years
group_by(zip,Year) %>%          # counting first time donors by postal code and selected years
summarise(nb_fixed_first_time_donors= n()) %>% #renaming
ungroup()  # ungrouping for safety reasons
```


## Number of repeated donors that donated to a fixed site per
```{r}
fix_preprocessed <- left_join(preprosessing ,firstevent,
by = c("zip", "Year")) %>%
  mutate( nb_fixed_repeat_donors= nb_fixed_donors_per_zip - nb_fixed_first_time_donors) 


# Joining all the data which has been made this far and counting the number of repeat donors (more than 1 donation) by postal code. Repeat donors equals number of donors per zip code minus number of first time donors per postal code.


fix_preprocessed$nb_fixed_first_time_donors[is.na(fix_preprocessed$nb_fixed_first_time_donors)] <- 0
fix_preprocessed$nb_fixed_repeat_donors[is.na(fix_preprocessed$nb_fixed_repeat_donors)] <- 0   # Deleting missing values
```



#joining the data with Paavodata
```{r}


processed_site_paavo <- paavodata %>%
rename(zip = pono, Year= vuosi) %>%  # Data needs to be translated to english from finnish
filter(Year == 2019) %>%   # Paavodata has been collected in the year 2017 so there is no need to get more years than one for our data. 
mutate(eligible_population = he_18_19 + he_20_24 + he_25_29 + he_30_34 + he_40_44 +
he_45_49 + he_50_54 + he_55_59 + he_60_64 + he_65_69) %>%   # Eligible population means age groups that can donate blood
dplyr::select(-Year)                                        # For losing two year variables
processed_site_paavo <-right_join(processed_site_paavo, fix_preprocessed, by = c("zip")) %>%          #joining with common factors
dplyr::select(zip, Year, eligible_population,hr_mtu, hr_ktu,ko_al_kork, ko_yl_kork, nb_fixed_donors_per_zip,nb_fixed_donations_per_zip, nb_fixed_first_time_donors, 
nb_fixed_repeat_donors, pt_tyott, pt_tyoll, ko_ika18y, hr_tuy,nimi) %>%          # Selecting wanted variables
rename (unemployed = pt_tyott,
employed = pt_tyoll,
medianincome= hr_mtu,
averageincome= hr_ktu,
population18= ko_ika18y,
bachelor_degree= ko_al_kork,
masters_degree= ko_yl_kork,
averageincome= hr_ktu,
medianincome= hr_mtu,
name= nimi) %>%          # mutating finnish variables to English.
mutate( prop_donors= nb_fixed_donors_per_zip/eligible_population,  #Proportion of donors is number of donors per postal code minus eligible population
nb_fixed_donation_per_act_donor= nb_fixed_donations_per_zip/nb_fixed_donors_per_zip, #  Number of donations per postal code divided by number of donors per postal code
prop_new_donors= nb_fixed_first_time_donors/eligible_population,   # Number of first time donors divided by eligible population per postal code
prop_repeat_donors= nb_fixed_repeat_donors/eligible_population,   #umber of repeat donors divided by eligible population per postal code
higher_education =bachelor_degree+masters_degree, # Combining university education together. 
proportion_inhabitants_with_higher_education= higher_education/eligible_population) # Higher education divided by eligible population per postal code


#Drop postal codes with no data
processed_site_paavo <- processed_site_paavo %>% filter(!is.na(eligible_population))


#Are there NA's left
processed_site_paavo[apply(processed_site_paavo,1,function(x){any(is.na(x))}),]

summary(processed_site_paavo)

```



## Make a variable which divides sites into mobile and fixed sites

```{r}

fixmob <- ilposdata %>% 
mutate(Year = year(dateonly)) %>%
filter(donat_phleb == "K") %>%
filter(Year == 2017| Year== 2018) 


colnames(fixmob )[grep("site",colnames(fixmob ))] <- "Site"
#Get the sites that are on the data
fixmob$Site <- recode(fixmob$Site,
                      H0091="Kivihaka",
                      H0092='Kivihaka',
                      H0093='Sanomatalo',
                      H0096='Espoo',
                      K0297='Kuopio',
                      L0564='Oulu',
                      T0179='Jyväskylä',
                      T0398='Lahti',
                      T0743='Seinäjoki',
                      T0837='Tampere',
                      U0853='Turku',
                      .default = levels(fixmob$Site)
                      )
#Add the mobile units in
fixmob$Site  <- as.character(fixmob$Site)
fixmob$Site[grep("^H1\\d{3}$",fixmob$Site ,perl = TRUE)] <- "Kivihaka mobile"
fixmob$Site [grep("^K1\\d{3}$",fixmob$Site ,perl = TRUE)] <- "Kuopio mobile"
fixmob$Site [grep("^L1\\d{3}$",fixmob$Site ,perl = TRUE)] <- "Oulu mobile"
fixmob$Site [grep("^S1\\d{3}$",fixmob$Site ,perl = TRUE)] <- "Seinäjoki mobile"
fixmob$Site [grep("^T1\\d{3}$",fixmob$Site ,perl = TRUE)] <- "Tampere mobile"
fixmob$Site [grep("^U1\\d{3}$",fixmob$Site ,perl = TRUE)] <- "Turku mobile"
#unknown places 
if (length(table(nchar(as.character(fixmob$Site)) == 0)) > 1 ) {
fixmob$Site [nchar(as.character(fixmob$Site )) == 0 ] <-  "unknown"}
# new variable 
fixmob$mobile <- recode(fixmob$Site,
                          Kivihaka="fixed site",
                          Sanomatalo="fixed site",
                          Espoo="fixed site",
                          Kuopio="fixed site",
                          Oulu="fixed site",
                          Jyväskylä="fixed site",
                          Lahti="fixed site",
                          Seinäjoki="fixed site",
                          Tampere="fixed site",
                          Turku="fixed site",
                          unknown="fixed site",
                          
                          "Kivihaka mobile"="mobile site",
                          "Kuopio mobile"="mobile site",
                          "Oulu mobile"="mobile site",
                          "Seinäjoki mobile"="mobile site",
                          "Tampere mobile"="mobile site",
                          "Turku mobile"="mobile site",.default = levels(fixmob$Site))


```



 #Compute distances between each zipcode and each mobile/fixed site. Then choose for each zipcode the closest fixed and the closest mobile.









## Get the number of donation opportunities for each mobile site in a years of interest

```{r}

```




for each zipcode compute proportion of donations made in fixed site.
```{r}

```





```{r}
#save(sitepreprosessed_paavo,
#file = paste0("name.",Sys.Date(),".RData"))
```

