---
title: "Visualising model data"
author: "Ilpo Arminen"
date: "8/16/2019"
output: github_document
---

```{r}
library(tidyverse)
library(lubridate)
library(knitr)
library(viridis)
library(table1)
```

```{r}
load("/home/ilpo/Paavo/data/model_data.RData")
```


```{r, results='asis}

label(model_data$medianincome)  <- "Median income of household per postal code"
label(model_data$averageincome) <- "Average income of household per postal code"
label(model_data$proportion_inhabitants_with_higher_education)  <- "Share of inhabitants with higher education per postal code"
label(model_data$prop_donors) <- "Proportion of donors per postal code "
label(model_data$nb_fixed_donations_per_zip) <- "Proportion of donations per postal code"
label(model_data$nb_fixed_donors_per_zip) <- "Number of donors per zip"
label(model_data$proportion_inhabitants_with_higher_education) <- "Proportion_inhabitants_with_higher_education"
label(model_data$nb_fixed_repeat_donors) <- "Number of donors who donated more than once per zip"
label(model_data$nb_fixed_first_time_donors) <- "Number of first time donors per zip"
label(model_data$nb_fixed_donation_per_act_donor) <- "Number of donations divided by numbers of donor per zip"
label(model_data$eligible_population) <- "Number of residents per zip who is suitable for donating blood"
units(model_data$medianincome) <- "Euro"
units(model_data$averageincome) <- "Euro" 

 table1(~prop_donors + eligible_population + nb_fixed_donors_per_zip + nb_fixed_donations_per_zip + nb_fixed_donors_per_zip + medianincome + averageincome + proportion_inhabitants_with_higher_education + proportion_inhabitants_with_higher_education + prop_repeat_donors + prop_new_donors + prop_donors + nb_fixed_donation_per_act_donor,data = model_data, overall = "Total", topclass="Rtable1-grid Rtable1-center")


```
```{r}
kable(model_data)
```

# Distribution of population of postal codes. 
```{r}
ggplot(data=model_data, mapping = aes(x= eligible_population))+
geom_histogram (binwidth= 1000)
```
### Shape of proportion of donors variable
```{r}
ggplot(data=model_data, mapping = aes(x= prop_donors))+
geom_histogram ()  
#scale_x_log10()
```
##  Change to logarithmic scale  helps  with the skewness.

```{r}
ggplot(data=model_data, mapping = aes(x= prop_donors))+
geom_histogram ()  +
scale_x_log10()
```
#
```{r}
ggplot(data= model_data, mapping= aes(x=minDist)) + 
geom_histogram()
```



#  Proportion of donors and median income
```{r}
ggplot(data=model_data, mapping = aes(x= medianincome, y= prop_donors))+
geom_point(mapping = aes(x= medianincome, y= prop_donors)) +
geom_smooth (method = "gam") + 
scale_x_log10() 
#facet_grid(Year~.)
```




## proportion of donors, median income and distance
```{r}
ggplot(data=model_data, mapping = aes(x= medianincome, y= prop_donors, color= minDist))+
geom_point(mapping = aes(x= medianincome, y= prop_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") +
facet_grid(Year~.)

```

###  proportion of donors, median income and population
```{r}
ggplot(data=model_data, mapping = aes(x= medianincome, y= prop_donors, color= eligible_population))+
geom_point(mapping = aes(x= medianincome, y= prop_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
facet_grid(Year~.)


```
####proportion of new donors plus median income +  (distance)


```{r}
ggplot(data=model_data, mapping = aes(x= medianincome, y= prop_new_donors, color= minDist))+
geom_point(mapping = aes(x= medianincome, y= prop_new_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
facet_grid(Year~.)

```
Proportion of  new donors, median income + (population)

```{r}
ggplot(data=model_data, mapping = aes(x= medianincome, y= prop_new_donors, color= eligible_population))+
geom_point(mapping = aes(x= medianincome, y= prop_new_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
facet_grid(Year~.)
```

##### proportion of  repeat donors, median income + (population)
```{r}
ggplot(data=model_data, mapping = aes(x= medianincome, y= prop_repeat_donors, color= eligible_population))+
geom_point(mapping = aes(x= medianincome, y= prop_repeat_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
facet_grid(Year~.)
```


##### Proportion of repeat donors, median income, colored by min distance
```{r}
ggplot(data=model_data, mapping = aes(x= medianincome, y= prop_repeat_donors, color= eligible_population))+
geom_point(mapping = aes(x= medianincome, y= prop_repeat_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
facet_grid(Year~.)
```

# proportion of donors and average income, colored by eligible population
```{r}
ggplot(data=model_data, mapping = aes(x= averageincome, y= prop_donors, color= eligible_population))+
geom_point(mapping = aes(x= averageincome, y= prop_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
facet_grid(Year~.)
```

##average income and distance 
```{r}
ggplot(data=model_data, mapping = aes(x= averageincome, y= prop_donors, color= minDist))+
geom_point(mapping = aes(x= averageincome, y= prop_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
facet_grid(Year~.)
```

####  proportion of new donors and average income (population)
```{r}
ggplot(data=model_data, mapping = aes(x= averageincome, y= prop_new_donors, color= eligible_population))+
geom_point(mapping = aes(x= averageincome, y= prop_new_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
facet_grid(Year~.)
```

Proportion of new donors  and average income (distance)
```{r}
ggplot(data=model_data, mapping = aes(x= averageincome, y= prop_new_donors, color= minDist))+
geom_point(mapping = aes(x= averageincome, y= prop_new_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
facet_grid(Year~.)
```



#####  proportion of repeat donors (population)
```{r}
ggplot(data=model_data, mapping = aes(x= averageincome, y= prop_repeat_donors, color= minDist))+
geom_point(mapping = aes(x= averageincome, y= prop_repeat_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
facet_grid(Year~.)
```

#Higher education and prop donor
```{r}
ggplot(data=model_data, mapping = aes(x= proportion_inhabitants_with_higher_education, y= prop_donors))+
geom_point(mapping = aes(x= proportion_inhabitants_with_higher_education,, y= prop_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
facet_grid(Year~.)
```

## Proportion of donors, higher education and distance
```{r}
ggplot(data=model_data, mapping = aes(x= proportion_inhabitants_with_higher_education, y= prop_donors, color= minDist))+
geom_point(mapping = aes(x= proportion_inhabitants_with_higher_education,, y= prop_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
facet_grid(Year~.)
```

### Propportion of donors, higher education and population
```{r}
ggplot(data=model_data, mapping = aes(x= proportion_inhabitants_with_higher_education, y= prop_donors, color= eligible_population))+
geom_point(mapping = aes(x= proportion_inhabitants_with_higher_education,, y= prop_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
facet_grid(Year~.)
```


##### proportion of new donors, higher education and eligible population
```{r}
ggplot(data=model_data, mapping = aes(x= proportion_inhabitants_with_higher_education, y= prop_new_donors, color= eligible_population))+
geom_point(mapping = aes(x= proportion_inhabitants_with_higher_education,, y= prop_new_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
facet_grid(Year~.)
```

####  proportion of repeat donors, higher education and distance 
```{r}
ggplot(data=model_data, mapping = aes(x= proportion_inhabitants_with_higher_education, y= prop_repeat_donors, color=minDist))+
geom_point(mapping = aes(x= proportion_inhabitants_with_higher_education,, y= prop_repeat_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
facet_grid(Year~.)
```


#####  proportion of repeat donors, higher education and eligible population
```{r}
ggplot(data=model_data, mapping = aes(x= proportion_inhabitants_with_higher_education, y= prop_repeat_donors, color=eligible_population))+
geom_point(mapping = aes(x= proportion_inhabitants_with_higher_education,, y= prop_repeat_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
facet_grid(Year~.)
```





















