---
title: "Visualising model data"
author: "Ilpo Arminen"
date: "8/16/2019"
output: github_document
---

```{r}
library(tidyverse)
library(lubridate)
library(knitr)
library(viridis)
library(table1)
library(GGally)
library(psycho)
```


```{r}
load("/home/ilpo/Paavo/data/model_data.RData")
```

#Tranforming variables and filtering the data

```{r}
model_data$minDistF<- cut(model_data$minDist,breaks = c(-1,1000, 5000,10000,20000,max(model_data$minDist))+1,include.lowest =TRUE,right = FALSE) #minDist looks weird in plots, so it needed to factorised. 

# new dataframe filtered for regression
modified_data <- model_data %>% 
filter(medianincome < 30000) %>%  # filtering few outliers away
mutate(minDistkm= minDist /  1000) %>% # meters to kilometers
filter(minDistkm > 0.5)        # filtering distances less than 500m 


```


#Table one from model data

```{r,results='asis'}

label(model_data$medianincome)  <- "Median income of household per postal code"
label(model_data$averageincome) <- "Average income of household per postal code"
label(model_data$proportion_inhabitants_with_higher_education)  <- "Share of inhabitants with higher education per postal code"
label(model_data$prop_donors) <- "Proportion of donors per postal code."
label(model_data$nb_fixed_donations_per_zip) <- "Proportion of donations per postal code"
label(model_data$nb_fixed_donors_per_zip) <- "Number of donors per zip"
label(model_data$proportion_inhabitants_with_higher_education) <- "Proportion_inhabitants_with_higher_education"
label(model_data$nb_fixed_repeat_donors) <- "Number of donors who donated more than once per zip"
label(model_data$nb_fixed_first_time_donors) <- "Number of first time donors per zip"
label(model_data$nb_fixed_donation_per_act_donor) <- "Number of donations divided by numbers of donor per zip"
label(model_data$eligible_population) <- "Number of residents per zip who is suitable for donating blood"
units(model_data$medianincome) <- "Euro"
units(model_data$averageincome) <- "Euro" 

 table1(~prop_donors + eligible_population + nb_fixed_donors_per_zip + nb_fixed_donations_per_zip + nb_fixed_donors_per_zip + medianincome + averageincome + proportion_inhabitants_with_higher_education + proportion_inhabitants_with_higher_education + prop_repeat_donors + prop_new_donors + prop_donors + nb_fixed_donation_per_act_donor,data = model_data, overall = "Total", topclass="Rtable1-grid Rtable1-center")


```

#Distributions of variables 

 ##Distribution of eligible population variable
```{r}
ggplot(data=model_data, mapping = aes(x= eligible_population))+
geom_histogram (binwidth= 500)

# Just to keep in mind that most of postal codes have around 2-5 thousand habitants. 
```
### Distribution of proportion of donors variable
```{r}
ggplot(data=model_data, mapping = aes(x= prop_donors))+
geom_histogram ()  

```
## Proportion of repeat donors 
```{r}
ggplot(data=model_data, mapping = aes(x= prop_new_donors))+
geom_histogram ()  

```

### Proportion of new donors 
```{r}
ggplot(data=model_data, mapping = aes(x= prop_repeat_donors))+
geom_histogram ()  

```


###Distribution of minimum distance variable
```{r}
ggplot(data= model_data, mapping= aes(x=minDist)) + 
geom_histogram()

# Most of postal codes in this data are around 10km distance to donation sites 
```
Distiribution of proportion_inhabitants_with_higher_education variable
```{r}
ggplot(data= model_data, mapping= aes(x=proportion_inhabitants_with_higher_education )) + 
geom_histogram()

# Proportion of higher 
```


#Median income


##Median income and proportion of donors
```{r}
ggplot(data=model_data, mapping = aes(x= medianincome, y= prop_donors))+
geom_point(mapping = aes(x= medianincome, y= prop_donors)) +
geom_smooth (method = "loess") + 
scale_x_log10() +
  labs(x = "Median income",
        y = "Proportion of donors",
        title = "Proportion of donors per median income of postal codes")
#facet_grid(Year~.)

#There seems to be quite a number of areas with high proportion of donors and lower than 20k median income. After checking this, it seems that areas are all from Lahti. Also need to pay attention that im not visualising the years separately, so  there are same data points with different year. it seems that there could be  (mild)  positive correlation between median income and proportion of donors, but only to some degree so that richest ares do not have highest prop donors and there are areas with high prop of donors and less than 20k median income, so there might be even negative correlation.  I would guess that Helsinki would fit to linear model better, than the whole dataset. 
```



## proportion of donors and median income with zip labels.

```{r}

 ggplot(data = model_data,
            mapping = aes(x = medianincome, y = prop_donors))+
geom_point(alpha= 0.25) +
scale_x_log10() + 
  geom_text(data = model_data ,check_overlap = TRUE, na.rm= TRUE, size=3,
mapping = aes(label = name)) +
 labs(x = "Median income",
        y = "Proportion of donors",
        title = "Proportion of donors per median income of postal codes")
#facet_grid(Year ~.)

# This if helpful for seeing that high proportion of donors seems to be areas near the fixed sites, atleast in Ruskeasuo, Etelä-Haaga and Lahti Asemanseutu and that Lahti has lot of areas where median income is low, but there are high proportion of donors.
```



## Median income,  distance with proportion of donors
```{r}
ggplot(data=model_data)+
geom_point(mapping = aes(x= medianincome, y= prop_donors,color= minDistF)) +
geom_smooth ( mapping = aes(x= medianincome, y= prop_donors)) + 
scale_color_viridis(discrete=TRUE,direction = -1) +
   labs(x = "Median income",
        y = "Proportion of donors",
        title = "Proportion of donors per median income and distance to donation site per postal code") +
facet_grid(Year~.)

# Distance does matter a lot as high proportion of donors comes from nearby fixed sites. 

```
##prop donors, median income and distance (km)
```{r}
ggplot(data=modified_data)+
geom_point(mapping = aes(x= medianincome, y= prop_donors,color= minDistkm)) +
geom_smooth ( mapping = aes(x= medianincome, y= prop_donors)) + 
scale_color_viridis(direction = -1) +
   labs(x = "Median income",
        y = "Proportion of donors",
        title = "Proportion of donors per median income and distance to donation site per postal code") +
facet_grid(Year~.)

# In this picture its pretty easy to see, that high proportion of donors comes from distances around 10 km
```


### Median income and population with  proportion of donors
```{r}
ggplot(data=model_data, mapping = aes(x= medianincome, y= prop_donors, color= eligible_population))+
geom_point(mapping = aes(x= medianincome, y= prop_donors)) +
geom_smooth (mapping = aes(x= medianincome, y= prop_donors)) + 
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
 labs(x = "Median income",
        y = "Proportion of donors",
        title = "Proportion of donors per median income and eligible population of postal codes")+ 
facet_grid(Year~.)

# Highest proportion of donors tend to come from smaller postal codes which sounds pretty reasonable, since there need to be smaller amount of donors per population. 
```

#proportion of new donors plus median income + distance

```{r}
ggplot(data=model_data)+
geom_point(mapping = aes(x= medianincome, y= prop_new_donors,color= minDistF)) +
geom_smooth(mapping = aes(x= medianincome, y= prop_new_donors))

# So this looks new donors and distance per median income. It seems that new donors do not live nearby of donation sites, so it looks pretty similar to all donors plot.  Distance tends to explain a lot more than median income, which seems to have negative trend. 

```
Proportion of  new donors, median income + population

```{r}
ggplot(data=model_data) +
geom_point(mapping = aes(x= medianincome, y= prop_new_donors,, color= eligible_population)) +
geom_smooth (aes(x= medianincome, y= prop_new_donors)) + 
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
 labs(x = "Median income",
        y = "Proportion of new donors",
        title = "Proportion of new donors per median income and eligible population of postal codes")+
facet_grid(Year~.)


# New donors also tend to come from postal codes from smaller population. 
```

# proportion of  repeat donors, median income + distance
```{r}
ggplot(data=model_data) +
geom_point(mapping = aes(x= medianincome, y= prop_repeat_donors, color=minDistF)) +
geom_smooth (mapping = aes(x= medianincome, y= prop_repeat_donors)) + 
#scale_color_viridis(discrete=FALSE,direction = -1) + 
   labs(x = "Median income",
        y = "Proportion of repeat donors",
        title = "Proportion of repeat donors per median income and distance to donation site per postal code") + 
facet_grid(Year~.)

#same trend is visible on repeat donors also. 
```


##### Proportion of repeat donors, median income, colored by eligible population
```{r}
ggplot(data=model_data, mapping = aes(x= medianincome, y= prop_repeat_donors, color= eligible_population))+
geom_point(mapping = aes(x= medianincome, y= prop_repeat_donors)) +
geom_smooth () + 
#scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
   labs(x = "Median income",
        y = "Proportion of  repeat donors",
        title = "Proportion of repeat donors per median income and eligible population per postal code") + 
facet_grid(Year~.)
```

# Average income 
##proportion of donors and average income, colored by eligible population
```{r}
ggplot(data=model_data, mapping = aes(x= averageincome, y= prop_donors, color= eligible_population))+
geom_point(mapping = aes(x= averageincome, y= prop_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
   labs(x = "Average income",
        y = "Proportion of donors",
        title = "Proportion of donors per average income and eligible population per postal code") +
facet_grid(Year~.)

# Averaga income and proportion of donors shows lack of donations from richest areas. Its reasonable that donations tend to come from areas where most of postal code tend to be. 
```

##average income and distance 
```{r}
ggplot(data=model_data)+
geom_point(mapping = aes(x= averageincome, y= prop_donors,color= minDistF)) +
geom_smooth (mapping = aes(x= averageincome, y= prop_donors)) + 
#scale_color_viridis() + 
   labs(x = "Average income",
        y = "Proportion of donors",
        title = "Proportion of donors per average income of postal codes") +
facet_grid(Year~.)

# Average income seems to have negative correlation with prop donors.  All the observations are scattered pretty randomly  into 20-30k area. 
```
#Proportion of new donors  and average income (distance)
```{r}
ggplot(data=model_data)+
geom_point(mapping = aes(x= averageincome, y= prop_new_donors,color= minDistF)) +
geom_smooth (mapping = aes(x= averageincome, y= prop_new_donors)) + 
 labs(x = "average income",
        y = "Proportion of donors",
        title = "Proportion of donors per average income and distance to donation site  per postal code") +
facet_grid(Year~.)
```



####  proportion of new donors and average income + population 
```{r}
ggplot(data=model_data, mapping = aes(x= averageincome, y= prop_new_donors, color= eligible_population))+
geom_point(mapping = aes(x= averageincome, y= prop_new_donors)) +
geom_smooth () + 
#scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
 labs(x = "Median income",
        y = "Proportion of  new donors",
        title = "Proportion of new donors per average income  and population size per postal code") +
facet_grid(Year~.)
```





#####  Proportion of repeat donors, average income and distance
```{r}
ggplot(data=model_data)+
geom_point(mapping = aes(x= averageincome, y= prop_repeat_donors,color= minDistF)) +
geom_smooth (mapping = aes(x= averageincome, y= prop_repeat_donors)) + 
#scale_x_log10() +
#scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
   labs(x = "Average income",
        y = "Proportion of repeat donors",
        title = "Proportion of repeat donors per average income and distance to donation site  of postal codes")+
facet_grid(Year~.)
```
##### Proportion of repeat donors, average income and population
```{r}
ggplot(data=model_data, mapping = aes(x= averageincome, y= prop_repeat_donors, color= eligible_population))+
geom_point(mapping = aes(x= averageincome, y= prop_repeat_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
   labs(x = "Average income",
        y = "Proportion of repeat donors",
        title = "Proportion of repeat donors per average income and eligible population of postal codes")+
facet_grid(Year~.)
```


#Higher education 

# Higher education and proportion of  donors
```{r}
ggplot(data=model_data, mapping = aes(x= proportion_inhabitants_with_higher_education, y= prop_donors))+
geom_point(mapping = aes(x= proportion_inhabitants_with_higher_education,, y= prop_donors)) +
geom_smooth (method="loess") + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
     labs(x = "Proportion of inhabitants with higher education per postal code",
        y = "Proportion of donors",
        title = "Proportion of donors and higher education per postal code") +
facet_grid(Year~.)

#  Median income and tertiary education plots seems to be scattered pretty similar and I would presume that if there is any correlation, it is weak linear one.  
```

##  Higher education and distance and proportion of donors
```{r}
ggplot(data=model_data) +
geom_point( mapping = aes(x= proportion_inhabitants_with_higher_education, y= prop_donors, color= minDistF)) +
geom_smooth ( mapping = aes(x= proportion_inhabitants_with_higher_education, y= prop_donors)) + 
#scale_x_log10() +
#scale_color_viridis(direction = -1) + 
   labs(x = "Proportion of inhabitants with higher education per postal code",
        y = "Proportion of donors",
        title = "Proportion of donors, higher education and distance to donation site per postal code") +  
facet_grid(Year~.)
#  Coloring the distance shows that donations do not favor educated areas, but areas with close distance to donation site. 
```


### Higher education and population and  proportion of donors
```{r}
ggplot(data=model_data, mapping = aes(x= proportion_inhabitants_with_higher_education, y= prop_donors, color= eligible_population)) +
geom_point(mapping = aes(x= proportion_inhabitants_with_higher_education,, y= prop_donors))  +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
   labs(x = "Proportion of inhabitants with higher education per postal code",
        y = "Proportion of donors",
        title = "Proportion of donors, higher education and eligible population  per postal code") +
facet_grid(Year~.)

# Population size of postal code tends to favor high number of donors in smaller places. This is in line with common sense since donations are measured by dividing the population size with number of donors.  Highest number of tertiary educated inhabitants behave similar to those highest number of median income areas, where there are quite a downfall in donation trend. Im not sure that the downfall is as drastic as it is with median income. 
```


##### proportion of new donors, higher education and eligible population
```{r}
ggplot(data=model_data, mapping = aes(x= proportion_inhabitants_with_higher_education, y= prop_new_donors, color= eligible_population))+
geom_point(mapping = aes(x= proportion_inhabitants_with_higher_education,, y= prop_new_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete = FALSE,direction = -1,trans="log") + 
   labs(x = "Proportion of inhabitants with higher education per postal code",
        y = "Proportion of new donors",
        title = "Proportion of new donors and higher education per postal code") +
facet_grid(Year~.)

# Eligible population shows that lot of high proportion of donor areas come from smaller postal codes even with new donors.  
```

####  proportion of repeat donors, higher education and distance 
```{r}
ggplot(data=model_data) +
geom_point(mapping = aes(x= proportion_inhabitants_with_higher_education, y= prop_repeat_donors, color=minDistF)) +
geom_smooth (mapping = aes(x= proportion_inhabitants_with_higher_education, y= prop_repeat_donors)) + 
     labs(x = "Proportion of inhabitants with higher education per postal code",
        y = "Proportion of repeat donors",
        title = "Proportion of repeat donors and higher education per postal code") +
facet_grid(Year~.)

# repeated donors shows the sane tremd than other plots with education. High proportion of donors come from close postal codes to fixed site. 
```
# 

#####  proportion of repeat donors, higher education and eligible population
```{r}
ggplot(data=model_data, mapping = aes(x= proportion_inhabitants_with_higher_education, y= prop_repeat_donors, color=eligible_population))+
geom_point(mapping = aes(x= proportion_inhabitants_with_higher_education,, y= prop_repeat_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
     labs(x = "Proportion of inhabitants with higher education per postal code",
        y = "Proportion of repeat  donors",
        title = "Proportion of repeat donors and higher education per postal code") +
facet_grid(Year~.)
```
 
# Distance 

#Distance and proportion of donors 

```{r}

ggplot(data=model_data)+
geom_point(mapping = aes(x= minDistF, y= prop_donors)) +
geom_smooth (mapping = aes(x= minDistF, y= prop_donors)) + 
#scale_x_log10() +
#scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
labs(x = "Distance to closest donation site",
        y = "Proportion of donors",
        title = "Proportion of  donors and distance to closest donation site") +
facet_grid(Year~.)


# 
```
```{r}
ggplot(data=model_data)+ # dataset where distance is kilometers and less than 500m distances are filtered out. 
geom_point(mapping = aes(x= minDist ,y= prop_donors)) +
geom_smooth (mapping = aes(x= minDist, y= prop_donors)) + 
scale_x_log10() +
#scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
labs(x = "Distance to closest donation site",
        y = "Proportion of donors",
        title = "Proportion of  donors and distance to closest donation site") +
facet_grid(Year~.)

# Distance seems to have moderate negative correlation with proportion of donors. 
```





## Distance and proportion of repeated donors
```{r}
ggplot(data=model_data)+
geom_point(mapping = aes(x= minDist, y= prop_repeat_donors)) +
geom_smooth ( mapping = aes(x= minDist, y= prop_repeat_donors)) + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
  labs(x = "Distance to closest donation site",
        y = "Proportion of repeat donors",
        title = "Proportion of  repeat donors and distance to closest donation site") +
facet_grid(Year~.)

# Same negative trend is visible on repeated donors also. 
```


### Distance and proportion of new donors

```{r}

ggplot(data=model_data, mapping = aes(x= minDist, y= prop_new_donors))+
geom_point(mapping = aes(x= minDist, y= prop_new_donors)) +
geom_smooth () + 
scale_x_log10() +
scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
    labs(x = "Distance to closest donation site",
        y = "Proportion of new donors",
        title = "Proportion of  new donors and distance to closest donation site") +
facet_grid(Year~.)

# New donors looks exactly the same as as other plots with proportion of donors and distance. 
```
# Median income and higher education

```{r}
ggplot(data=model_data)+
geom_point(mapping =aes( x= medianincome, y= proportion_inhabitants_with_higher_education)) +
geom_smooth ( mapping = aes(x= medianincome, y= proportion_inhabitants_with_higher_education, method= "loess"))

#Median income and higher education tends to correlate with strongly. 
```

# Distance in kilometers with filtered data
```{r}
ggplot(data=modified_data)+ # dataset where distance is kilometers and less than 500m distances are filtered out. 
geom_point(mapping = aes(x= minDistkm, y= prop_donors)) +  # meters turned into kilometers. 
geom_smooth (mapping = aes(x= minDistkm, y= prop_donors)) + 
#scale_x_log10() +
#scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
labs(x = "Distance to closest donation site",
        y = "Proportion of donors",
        title = "Proportion of  donors and distance to closest donation site") +
facet_grid(Year~.)

# Distance seems to have moderate negative correlation with proportion of donors. 
```
```{r}
ggplot(data=modified_data)+ # dataset where distance is kilometers and less than 500m distances are filtered out. 
geom_point(mapping = aes(x= minDistkm, y= prop_repeat_donors)) +  # meters turned into kilometers. 
geom_smooth (mapping = aes(x= minDistkm, y= prop_repeat_donors)) + 
scale_x_log10() +
#scale_color_viridis(discrete=FALSE,direction = -1,trans="log") + 
labs(x = "Distance to closest donation site",
        y = "Proportion of donors",
        title = "Proportion of  donors and distance to closest donation site") +
facet_grid(Year~.)

```


#Correlation matrix and plot
```{r}
cordata <- model_data %>% # Making a new dataframe so i dont count correlation to all variables. 
dplyr::select(prop_donors,prop_new_donors,prop_repeat_donors, medianincome, averageincome,minDist,proportion_inhabitants_with_higher_education, eligible_population) #selecting wanted variables


cor <- psycho::correlation(cordata) # R doesn't recognize this particular correlation function without package name first 
plot(cor)#visualise the correlation matrix 
#print(cor)

```


```{r,results='asis'}
summary(cor) # prints correlation matrix 
# Have to check the multicollinearity later as higher education and median income have high correlation (0.77). 
write_csv(summary(cor), "myformattedcortable.csv")
```



```{r}
save(modified_data, #saving the filtered data for regression
file = paste0("name.",Sys.Date(),".RData"))
```





